---
title: "hw1"
author: "David DeStephano"
date: "February 25, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RNHANES)
library(tidyverse)
library(summarytools)
library(stargazer)
library(leaps)
library(caret)
library(ModelMetrics)
library(pls)
```

#Import the data
```{r}
test<-read_csv("solubility_test.csv")

train<-read_csv("solubility_train.csv")
```


#Linear Model
```{r}
fit_lm <-lm(Solubility~., data = train)
pred_lm <-predict(fit_lm, test)
mse(test$Solubility, pred_lm)
```

#Preprocess data for ML
```{r}
x <-model.matrix(Solubility~.,train)[,-1]
y <- train$Solubility

# test data
x2 <-model.matrix(Solubility~.,test)[,-1]
y2 <- test$Solubility
```

```{r}
#
xTrans <- preProcess(train)
#trainDescr <- predict(xTrans, train)
#xTrans <- preProcess(trainDescr)
```
All variables are already centered and scaled

#ctrl1 settings
```{r}
ctrl1 <-trainControl(method = "repeatedcv", number = 10, repeats = 5)

set.seed(2)
```

#Linear model using Caret
```{r}
lm.fit <-train(x, y,method = "lm",trControl = ctrl1)

predy2.lm <-predict(lm.fit, newdata = x2)

mse(y2, predy2.lm)

```

#Ridge regression
```{r}
ridge.fit <-train(x, y,
                  method = "glmnet",
                  tuneGrid =expand.grid(alpha = 0,
                                        lambda=exp(seq(-1, 10, length=100))),
                  #preProc = c("center","scale"),
                  trControl = ctrl1)

predy2.ridge <-predict(ridge.fit, newdata = x2)

mse(y2, predy2.ridge)
```

##Checking paramters and best lambda value
```{r}
plot(ridge.fit, xTrans =function(x)log(x))

ridge.fit$bestTune

#coef(ridge.fit$finalModel,ridge.fit$bestTune$lambda)
```
#Lasso regression
```{r}
lasso.fit <-train(x, y,
                  method = "glmnet",
                  tuneGrid =expand.grid(alpha = 1,
                                      lambda=exp(seq(-1, 10, length=100))),
                  #preProc = c("center","scale"),
                  trControl = ctrl1)

predy2.lasso <-predict(lasso.fit, newdata = x2)

mse(y2, predy2.lasso)
```

##Checking paramters and best lambda value
```{r}
plot(lasso.fit, xTrans =function(x) log(x))

lasso.fit$bestTune

coef(lasso.fit$finalModel,lasso.fit$bestTune$lambda)
```

The test error is 1.328538, and there were 8 variables with non-zero coefficient estimates

#Principal component regression
```{r}
pcr.fit <-train(x, y,
                method = "pcr",
                tuneGrid  =data.frame(ncomp = 1:228),
                trControl = ctrl1,
                #preProc =c("center", "scale"))


predy2.pcr4 <-predict(pcr.fit, newdata = x2)
mse(y2, predy2.pcr4)
```

The test error is 0.5489785 and M is 160

```{r}
ggplot(pcr.fit, highlight = TRUE)+ theme_bw()
pcr.fit$bestTune
```






#(e) Briefy discuss the results obtained in (a)-(d).

```{r}
resamp <- resamples(list(lm=lm.fit,
                         ridge=ridge.fit,
                         lasso=lasso.fit,
                         pcr=pcr.fit))

summary(resamp)
```


#(f) Which model will you choose for predicting solubility?
I would chose the ridge regression model as it has the lowest MSE




#######
#How do i use training data to find best model?
#Also validation set is not cross validation...
